---
title: "Assignment 5"
author: "Lokesh Arora, Ankita Shinde"
date: "10/8/2020"
output: pdf_document
Github Link: https://github.com/ankita1598/Walmart
---
```{r}
library(lubridate)
library("plyr")
library("ggplot2")
library(RColorBrewer)
library("dplyr")
```

```{r}
dataset = read.csv("data.csv", header= T)
head(dataset)
```

We can see that there are few null values in the data set for column Markdown 1 - 5. We will also split the data column in 3 as Day, Month and Year.

```{r}
dataset$Year <- year(ymd(dataset$Date))
dataset$Month <- month(ymd(dataset$Date)) 
dataset$Day <- day(ymd(dataset$Date))
dataset$Dept = as.factor(dataset$Dept)
dataset$Store = as.factor(dataset$Store)
dataset$MarkDown1[is.na(dataset$MarkDown1)] = 0
dataset$MarkDown2[is.na(dataset$MarkDown2)] = 0
dataset$MarkDown3[is.na(dataset$MarkDown3)] = 0
dataset$MarkDown4[is.na(dataset$MarkDown4)] = 0
dataset$MarkDown5[is.na(dataset$MarkDown5)] = 0
dataset = fastDummies::dummy_cols(dataset, select_columns = "Type")
dataset$IsHoliday[dataset$isHoliday == "False"] = 0
dataset$IsHoliday[dataset$isHoliday == "True"] = 1
head(dataset)
```
```{r}
dim(dataset)
```

As we can see from the result now there are 19 columns and 421570 rows.

```{r}
names(dataset)
```

Here are all the columns.

```{r}
dataset$Dept = as.numeric(as.factor(dataset$Dept))
dataset$Store = as.numeric(as.factor(dataset$Store))
features = c("Store","Dept","IsHoliday","Type_A","Type_B","Type_C","Size","Temperature","Fuel_Price","MarkDown1","MarkDown2","MarkDown3","MarkDown4","MarkDown5","CPI","Unemployment","Year","Month","Day")
correlation = cor(select(dataset,features))
correlation
```
```{r}
dataset_pca = prcomp(select(dataset,features), scale=TRUE)
dataset_pca
```

```{r}
summary(dataset_pca)
```
```{r}
dataset_eigen = dataset_pca$sdev^2
names(dataset_eigen) <- paste("PC",1:19,sep="")
dataset_eigen

```
```{r}
sumlambdas = sum(dataset_eigen)
sumlambdas
```
```{r}
propvar = dataset_eigen/sumlambdas
propvar
```
```{r}
dataset_cumvar <- cumsum(propvar)
dataset_cumvar
```
```{r}
matlambdas <- rbind(dataset_eigen,propvar,dataset_cumvar)
rownames(matlambdas) <- c("Eigenvalues","Prop. variance","Cum. prop. variance")
round(matlambdas,4)
```
```{r}
summary(dataset_pca)
```
```{r}
dataset_pca$rotation
```
```{r}
print(dataset_pca)
```

1st Option
Based on retating components that account for 70% to 90% of the variance, we need to retain PC1 to PC8 or PC1 to PC12.

2nd Option
Based on the rule of sum to choose all components with eigen values larger than 0.7, we need to retain PC1 to PC12.


```{r}
dataset_pca$x
```
```{r}
weeklySales <- data.frame(WeeklySales=dataset$weeklySales)
dataset2_pca <- cbind(weeklySales, dataset_pca$x)
dataset2_pca
```
```{r}
tabmeansPC <- aggregate(dataset2_pca[,2:13],by=list(weeklySales=dataset$weeklySales),mean)
tabmeansPC
```
```{r}
tabmeansPC <- tabmeansPC[rev(order(tabmeansPC$weeklySales)),]
tabmeansPC
```
```{r message=TRUE, warning=TRUE}
tabfmeans <- t(tabmeansPC[,-1])
```

```{r}

colnames(tabfmeans) <- t(as.vector(tabmeansPC[1]))
```
```{r}
tabsdsPC <- aggregate(dataset2_pca[,2:13],by=list(weeklySales=dataset$weeklySales),sd)
tabfsds <- t(tabsdsPC[,-1])
colnames(tabfsds) <- t(as.vector(tabsdsPC[1]))
```
```{r}
plot(dataset_eigen, xlab = "Component number", ylab = "Component variance", type = "l", main = "Scree diagram")
```

```{r}
plot(log(dataset_eigen), xlab = "Component number",ylab = "log(Component variance)", type="l",main = "Log(eigenvalue) diagram")
```
```{r}
summary(dataset_pca)
```
```{r}
cov(dataset_pca$x)
```
```{r}
dataset_pca$rotation[,1]
dataset_pca$rotation
```
```{r}
#plot(select(dataset,features))
```
```{r}
center <- dataset_pca$center
scale <- dataset_pca$scale
new_dataset <- as.matrix(select(dataset,features))
new_dataset
```

```{r}
drop(scale(new_dataset,center=center, scale=scale)%*%dataset_pca$rotation[,1])
```
```{r}
predict(dataset_pca)[,1]
```
```{r}
out <- sapply(1:13, function(i){plot(dataset$weeklySales,dataset_pca$x[,i],xlab=paste("PC",i,sep=""),ylab="Weekly Sales")})
```

```{r}
#pairs(dataset_pca$x[,1:13], ylim = c(-6,4),xlim = c(-6,4),panel=function(x,y,...){text(x,y,dataset$weeklySales)})
```




